---
title: "ToothGrowthAnalyses"
author: "AlekseyVosk"
date: '3 June 2018'
output: html_document
---

For research we will need functions from external packages.
```{r}
library(tidyverse)
```

## Data and exploratory data analyses
Dataset contanes 60 observations on 3 variables:  
1. len - Tooth length  
2. supp - Supplement type: *VC* (ascorbic acid) or *OJ* (orange juice)  
3. dose - Dose in milligrams/day  

Loading data as tibble and looking at data's structure.
```{r}
dataset <- as_tibble(ToothGrowth)
str(dataset)
```
Let's compare tooth length in groups with different doses and different supplement types.
```{r}
dataset_f <- mutate(dataset, dose = as.factor(dose))
table(dataset_f$supp, dataset_f$dose)
group_by(dataset_f, dose) %>%
    summarise(len = mean(len))
group_by(dataset_f, supp) %>%
    summarise(len = mean(len))
```
Tooth length differs between each groups, to visualise it, let's make boxplots:
```{r}
g <- ggplot(data = dataset_f, mapping = aes(y = len)) +
    labs(y = "Tooth length", 
         title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs")
g + geom_boxplot(mapping = aes(x = dose)) +
    labs(x = "Dose of supplement , milligrams/day ",
         title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs, by dose")
g + geom_boxplot(mapping = aes(x = supp)) +
    labs(x = "Supplement type, OJ (orange juice) or VC (ascorbic acid)",
         title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs, by supplement")
g + geom_boxplot(mapping = aes(x = dose)) +
    facet_wrap(~ supp) +
    labs(x = "Dose of supplement , milligrams/day ",
         title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs, by dose and supplement")
```

Exploratory analyses shows us, than the tooth growth in guinea pigs differs in groups with different doses of vitamin C, but not so much in groups with different supplement type.