---
title: "ToothGrowthAnalyses"
author: "AlekseyVosk"
date: '3 June 2018'
output: html_document
---

For research we will need functions from external packages.
```{r}
library(tidyverse)
```

## Data and exploratory data analyses
Dataset contanes 60 observations on 3 variables:  
1. len - Tooth length  
2. supp - Supplement type: *VC* (ascorbic acid) or *OJ* (orange juice)  
3. dose - Dose in milligrams/day  

Loading data as tibble and looking on head/tail of data and it's structure.
```{r}
dataset <- as_tibble(ToothGrowth)
head(dataset, 3)
tail(dataset, 3)
str(dataset)
```
To see how data is distributed, let's make table with 2 categorical variables (supp and dose) and boxplot of continuous variable (len).
```{r}
dataset_f <- mutate(dataset, dose = as.factor(dose))
table(dataset_f$supp, dataset_f$dose)
g <- ggplot(data = dataset_f) +
    geom_boxplot(mapping = aes(x = dose, y = len)) +
    labs(x = "Dose of supplement , milligrams/day ", y = "Tooth length", 
         title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs")
g
g + facet_wrap(~ supp) +
    labs(title = "Effect of Vitamin C on Tooth Growth in Guinea Pigs, by Supplement type")
```